---
title: "Simulation Studies"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Fake data simulation plays an important role in determining how well different model selection methods work in practice and complement Cross-Validation.  Because we know the true data mechanism,  we can look at how well various procedures can capture the "truth".

We will generate data according to a model from a simulation study in a paper by Nott & Kohn.

First let's set our random seed in case we need to rerun parts later.

```{r jenny, echo=TRUE}
# set the random seed so that we can replicate results.
set.seed(8675309)
```

### True parameters

In order to simulate data, we need to specify the values of the  "true" parameters.  For this study we will use

```{r true}
# true parameters
sigma = 2.5
betatrue = c(4,2,0,0,0,-1,0,1.5, 0,0,0,1,0,.5,0,0,0,0,-1,1,4)
#          int|    X1                            | X2     |X3          
```



### Generate Data

We are now going to generate an 100  X matrices with correlated columns
```{r datasets} 
#sample size
n = 50

# part of dataframe name
fname=rep("df",100)

# create 100 datasets
for (i in 1:100) {
  
# generate some satandard normals
  Z = matrix(rnorm(n*10, 0, 1), ncol=10, nrow=n)
  
#  Create X1 by taking linear cominations of Z to induce correlation among X1 components
  
  X1 = cbind(Z, (Z[,1:5] %*% c(.3, .5, .7, .9, 1.1) %*% t(rep(1,5)) +
             matrix(rnorm(n*5, 0, 1), ncol=5, nrow=n)))
# generate X2 as a standard normal  
  X2 <- matrix(rnorm(n*4,0,1), ncol=4, nrow=n)
  
# Generate X3 as a linear combination of X2 and noise  
  X3 <- X2[,4]+rnorm(n,0,sd=0.1)
  
# combine them  
  X <- cbind(X1,X2,X3)
  
# subtract off the column means
  X = sweep(X, 2, apply(X,2, mean)) 

# Generate mu     
# X does not have a column of ones for the intercept so need to add the intercept  
# for true mu  
  mu = betatrue[1] + X %*% betatrue[-1] 
  
# now generate Y  
  Y = mu + rnorm(n,0,sigma)  
  
# make a dataframe and save it
  df = data.frame(Y, X, mu)
  fname[i] = paste("df", as.character(i), sep="")
  save(df, file=fname[i])
}
```



### Estimation

Let's see how well we do at estimating the true $\beta$s

```{r MSE-OLS}
MSE.OLS  = rep(0,100)

for( i in 1:100) {
  rm(df)
  load(fname[i])
  nk.ols = lm(Y ~ . -mu, data=df)
  coef.ols = coef(nk.ols)
  MSE.OLS[i] = sum((betatrue - coef.ols)^2)
#  print(c(i, MSE.OLS[i]))
}

#  mean squared error loss for estimating beta
mean(MSE.OLS)
```

While a single number summary is great, how much variation is there across simulations?  What is the worse behaviour?

```{r}
hist(MSE.OLS)
```

### Other estimators

Besides OLS for the "full model" we could use stepwise selection or other methods or look at other loss functions for estimating the mean or prediction.

## To do individually/teams  (will be part of HW)

1.  What happens if you change the "loss function" so instead of  MSE for $\beta$ it is MSE for estimating $\mu = X \beta$  with OLS?  (write a function for evaluating this loss)

2. Use stepwise selection with AIC and BIC for the simulated data and evaluate the loss for estimating $\beta$ and estimating $\mu$

3.  What if the loss function is 0 if you do not pick the true model but -1 if you do choose the correct model.
Write a function to evaluate this loss and then use this with different model selection methods (i.e stepwise wit AIC and  BIC)








